---
title: "Unsupervised Learning"
output: github_document
---

## How to get set up for this workshop

### Our virtual space:

[http://backchannelchat.com/chat/parmw](http://backchannelchat.com/chat/parmw)

### Getting this document on your computer:

1. Go to the GitHub repository here: [https://github.com/davidklaing/data_science_workshops](https://github.com/davidklaing/data_science_workshops)
2. Click the green button on the right that says "Clone or download".
3. Click "Download ZIP". (If you're proficient with git, feel free to clone the repository.)
4. Create a folder on your computer to store your work, and store your ZIP file there.
5. Double-click your ZIP file to unzip it and get all the code.

### Getting R and Rstudio

1. Download and install R from here: [http://cran.stat.sfu.ca/](http://cran.stat.sfu.ca/).
2. Download and install RStudio Desktop (Open Source Edition) from here: [https://www.rstudio.com/products/rstudio/#Desktop](https://www.rstudio.com/products/rstudio/#Desktop).

### Getting ready to play!

1. In RStudio, open `unsupervised_learning.Rmd`, a file in `YOUR_FOLDER/data_science_workshops/unsupervised_learning/src/`. (That's this file!)
2. In the code snippet below, remove the hashtags from both lines, and click the green "play" button on the right to install `dplyr` and `ggplot2`, the two packages you'll need for the workshop.

```{r}
#install.packages("dplyr")
#install.packages("ggplot2")

library(dplyr)
library(ggplot2)
library(purrr)
```

### Having installation problems?

Option 1: Ask for help! We have volunteers who can help troubleshoot.

Option 2a: Find a partner and follow along together — most of the exercises can be done collaboratively. Your installation problem is almost certainly solvable — we just might not have time today.

Option 2b: Use [this datacamp light](https://cdn.datacamp.com/dcl/standalone-example.html) page to write and execute code in your browser. (A couple caveats: doesn't come with the titanic data, and might end up distracting you from the lesson. If this is your first time programming, I recommend option 2a.)

## Rstudio review

Executing code: move your cursor to the line you want to execute, and hit CTRL+ENTER

```{r}
print("hello!")
1+1
```

Looking up documentation about a function or object: prepend the name of the object with a question mark, and execute that line.

```{r}
print
```

```{r}
set.seed(3)

generate_height <- function(species) {
  if (species == "wizard") {
    max(0, rnorm(1, 1.8, 0.3) + rnorm(1, 0, 0.4))
  } else if (species == "dwarf") {
    max(0, rnorm(1, 1.2, 0.2) + rnorm(1, 0, 0.4))
  } else if (species == "giant") {
    max(0, rnorm(1, 2.6, 0.5) + rnorm(1, 0, 0.4))
  }
}

generate_magical_powers <- function(species) {
  if (species == "wizard") {
    max(0, rnorm(1, 8, 1.2))
  } else if (species == "dwarf") {
    max(0, rnorm(1, 3.5, 0.6))
  } else if (species == "giant") {
    max(0, rnorm(1, 1, 0.7))
  }
}

magical_data <- data_frame(
  species = sample(c("wizard", "dwarf", "giant"), 100, replace = TRUE)
) %>% 
  mutate(
    height = map_dbl(species, generate_height),
    magical_powers = map_dbl(species, generate_magical_powers)
  )

unsupervised_magical_data <- magical_data %>% 
  select(height, magical_powers)

magical_data %>% 
  ggplot() +
  geom_point(
    aes(
      x = height,
      y = magical_powers,
      color = species
    )
  )

unsupervised_magical_data %>% 
  ggplot() +
  geom_point(
    aes(
      x = height,
      y = magical_powers
    )
  )
```

```{r}
?kmeans

kmeans_result <- kmeans(select(magical_data, -species), 3)

kmeans_result$cluster

magical_data %>% 
  mutate(cluster = kmeans_result$cluster) %>% 
  ggplot() +
  geom_point(
    aes(
      x = height,
      y = magical_powers,
      color = factor(species)
    )
  )
```

# Implement k-means!

Steps:

1. Choose k
2. Initialize k random centroids (can choose points)
3. Compute the distance from each point to the k centroids
4. Assign each point to a group corresponding to its closest centroid
5. Repeat steps 3 and 4 until the points no longer change groups.
