---
title: "Applied Statistics for Data Science"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How to get set up for this workshop

### Our virtual space:

[https://todaysmeet.com/wwc_stats](https://todaysmeet.com/wwc_stats)

### Getting this document on your computer:

1. Go to the GitHub repository here: [https://github.com/davidklaing/data_science_workshops](https://github.com/davidklaing/data_science_workshops)
2. Click the green button on the right that says "Clone or download".
3. Click "Download ZIP". (If you're proficient with git, feel free to clone the repository.)
4. Create a folder on your computer to store your work, and store your ZIP file there.
5. Double-click your ZIP file to unzip it and get all the code.

### Getting R and Rstudio

1. Download and install R from here: [http://cran.stat.sfu.ca/](http://cran.stat.sfu.ca/).
2. Download and install RStudio Desktop (Open Source Edition) from here: [https://www.rstudio.com/products/rstudio/#Desktop](https://www.rstudio.com/products/rstudio/#Desktop).

### Getting ready to play!

1. In RStudio, open `applied_statistics.Rmd`, a file in `YOUR_FOLDER/data_science_workshops/applied_statistics`. (That's this file!)
2. In the code snippet below, remove the hashtags from both lines, and click the green "play" button on the right to install `dplyr`, `ggplot2`, and `titanic`, the three packages you'll need for the workshop.

```{r}
#install.packages("dplyr")
#install.packages("ggplot2")
#install.packages("titanic")

library(dplyr)
library(ggplot2)
```

### Having installation problems?

Option 1: Ask for help! We have volunteers who can help troubleshoot.

Option 2a: Find a partner and follow along together — most of the exercises can be done collaboratively. Your installation problem is almost certainly solvable — we just might not have time today.

Option 2b: Use [this datacamp light](https://cdn.datacamp.com/dcl/standalone-example.html) page to write and execute code in your browser. (A couple caveats: doesn't come with the titanic data, and might end up distracting you from the lesson. If this is your first time programming, I recommend option 2a.)

## A challenge of inference

Let's generate some data. Don't worry yet about how it's being generated — all you need to know is that there's a random process that determines which numbers are most likely, and I'm defining that random process with this function.

```{r}
my_sample_df <- data_frame(
  some_variable = rpois(n = 100, lambda = 4)
)
```

Let's take a peak at this sample we've generated.

```{r}
my_sample_df
```

Our first insight: even with a sample as small as this, it's pretty hard to summarize what we've learned from this data. 

```{r}
t.test(my_sample_df$some_variable, alternative = "greater", mu = 3)
```


```{r}
my_sample_df %>% 
  ggplot() +
  geom_histogram(aes(x = some_variable), binwidth = 1)
```













