---
title: "Supervised Learning"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How to get set up for this workshop

### Our virtual space:

[https://todaysmeet.com/wwc_stats](https://todaysmeet.com/wwc_stats)****FIX

### Getting this document on your computer:

1. Go to the GitHub repository here: [https://github.com/davidklaing/data_science_workshops](https://github.com/davidklaing/data_science_workshops)
2. Click the green button on the right that says "Clone or download".
3. Click "Download ZIP". (If you're proficient with git, feel free to clone the repository.)
4. Create a folder on your computer to store your work, and store your ZIP file there.
5. Double-click your ZIP file to unzip it and get all the code.

### Getting R and Rstudio

1. Download and install R from here: [http://cran.stat.sfu.ca/](http://cran.stat.sfu.ca/).
2. Download and install RStudio Desktop (Open Source Edition) from here: [https://www.rstudio.com/products/rstudio/#Desktop](https://www.rstudio.com/products/rstudio/#Desktop).

### Getting ready to play!

1. In RStudio, open `supervised_learning.Rmd`, a file in `YOUR_FOLDER/data_science_workshops/supervised_learning/src/`. (That's this file!)
2. In the code snippet below, remove the hashtags from both lines, and click the green "play" button on the right to install `dplyr`, `ggplot2`, and `titanic`, the three packages you'll need for the workshop.

```{r}
#install.packages("dplyr")
#install.packages("ggplot2")
#install.packages("titanic")

library(dplyr)
library(ggplot2)
library(titanic)
```

### Having installation problems?

Option 1: Ask for help! We have volunteers who can help troubleshoot.

Option 2a: Find a partner and follow along together — most of the exercises can be done collaboratively. Your installation problem is almost certainly solvable — we just might not have time today.

Option 2b: Use [this datacamp light](https://cdn.datacamp.com/dcl/standalone-example.html) page to write and execute code in your browser. (A couple caveats: doesn't come with the titanic data, and might end up distracting you from the lesson. If this is your first time programming, I recommend option 2a.)

## Warm-up exercise (in pairs)

Based on the graphs below, how would you predict future values of $y$ knowing only the corresponding values of $x$?

```{r}
df1 <- data_frame(
  x = c(0,1.4,2,3.7,4,5.1,6,7.5,8.4,9.7),
  y = c(1,1,1,1,1,3,3,3,3,3)
)

df1 %>% 
  ggplot() +
  geom_point(
    aes(x = x, y = y)
  ) +
  scale_y_continuous(
    limits = c(0,4),
    breaks = c(0,1,2,3,4)
  ) +
  scale_x_continuous(
    breaks = c(0,1,2,3,4,5,6,7,8,9,10)
  ) +
  labs(
    title = "Graph 1"
  )
```

```{r}
df2 <- data_frame(
  x = c(0,1.1,2,3.3,4.4,5,6.2,7.7,8.1,9),
  y = c(0,2.2,4,6.6,8.8,10,12.4,15.2,16.2,18)
)

df2 %>% 
  ggplot() +
  geom_point(
    aes(x = x, y = y)
  ) +
  scale_y_continuous(
    breaks = seq(0,20,2)
  ) +
  scale_x_continuous(
    breaks = c(0,1,2,3,4,5,6,7,8,9,10)
  ) +
  labs(
    title = "Graph 2"
  )
```

```{r}
df2 <- data_frame(
  x = c(0,1.1,2,3.3,4.4,5,6.2,7.7,8.1,9),
  y = c(1.1,1.3,0.9,0.87,8.8,8.4,7.4,6,6.4,6.9)
)

df2 %>% 
  ggplot() +
  geom_point(
    aes(x = x, y = y)
  ) +
  scale_y_continuous(
    breaks = seq(0,20,2)
  ) +
  scale_x_continuous(
    breaks = c(0,1,2,3,4,5,6,7,8,9,10)
  ) +
  labs(
    title = "Graph 3"
  )
```

## Supervised learning

What it is

Why it's useful

How to do it


