---
title: "Applied Statistics for Data Science"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## How to get set up for this workshop

### Our virtual space:

[https://todaysmeet.com/wwc_stats](https://todaysmeet.com/wwc_stats)

### Getting this document on your computer:

1. Go to the GitHub repository here: [https://github.com/davidklaing/data_science_workshops](https://github.com/davidklaing/data_science_workshops)
2. Click the green button on the right that says "Clone or download".
3. Click "Download ZIP". (If you're proficient with git, feel free to clone the repository.)
4. Create a folder on your computer to store your work, and store your ZIP file there.
5. Double-click your ZIP file to unzip it and get all the code.

### Getting R and Rstudio

1. Download and install R from here: [http://cran.stat.sfu.ca/](http://cran.stat.sfu.ca/).
2. Download and install RStudio Desktop (Open Source Edition) from here: [https://www.rstudio.com/products/rstudio/#Desktop](https://www.rstudio.com/products/rstudio/#Desktop).

### Getting ready to play!

1. In RStudio, open `applied_statistics.Rmd`, a file in `YOUR_FOLDER/data_science_workshops/applied_statistics`. (That's this file!)
2. In the code snippet below, remove the hashtags from both lines, and click the green "play" button on the right to install `dplyr`, `ggplot2`, and `titanic`, the three packages you'll need for the workshop.

```{r}
#install.packages("dplyr")
#install.packages("ggplot2")
#install.packages("titanic")

library(dplyr)
library(ggplot2)
```

### Having installation problems?

Option 1: Ask for help! We have volunteers who can help troubleshoot.

Option 2a: Find a partner and follow along together — most of the exercises can be done collaboratively. Your installation problem is almost certainly solvable — we just might not have time today.

Option 2b: Use [this datacamp light](https://cdn.datacamp.com/dcl/standalone-example.html) page to write and execute code in your browser. (A couple caveats: doesn't come with the titanic data, and might end up distracting you from the lesson. If this is your first time programming, I recommend option 2a.)

## Warm-up exercise — in pairs

The code snippet below reads in a CSV file with just one column — a vector of numbers. Recalling what you may have learned in the previous workshop, try visualizing these numbers to get a sense of their distribution. To consider:

- What type(s) of plots might be appropriate in this case? 
- If you had to summarize the sample numerically, what would you do?
- Say we draw one more number from the same process that generated this sample — what would you predict would be its value?
- Say we draw 100,000 more numbers from the same process that generated this sample — what would you predict about their values?

```{r}
sample_df <- readr::read_csv("../data/sample_df.csv")

# Write your code here!
```

## The most common statistical tasks in data science

1. Understanding the distribution of a sample.
2. Summarizing information about a sample.
3. Predicting the future.
4. Characterizing the unknown.

## 1. Understanding the distribution of a sample

Key concepts:

- Histograms
- Density plots
- Boxplots
- Bar plots

Useful R functions:

- `ggplot()`, `geom_histogram()`, `geom_density()`, `geom_boxplot()` `geom_bar()`

Common ways to fool yourself:

- Viewing your distribution at the wrong level of aggregation
- Overlaying group distributions on an unaligned scale 

## 2. Summarizing information about a sample

Key concepts:

- Measures of location — mean, median, (less commonly) mode
- Measures of spread — standard deviation, variance, maximum, minimum, percentiles

Useful R functions:

- `summary()`, `mean()`, `median()`, `sd()`, `var()`, `max()`, `min()`, `quantile()`

## 3. Predicting the future

Key concepts:

- Expected value, error minimization

Useful R functions:

- `mean()`

...more to come next month, when we discuss machine learning!

## 4. Characterizing the unknown (a.k.a. *inference*)

Key concepts:

- Sample, population, maximum likelihood estimation, hypothesis test, bootstrap

Useful R functions:

- `t.test()`





